Rectangulotopes

### 2024-04-04 draft/rectangulotopes.tex
Sections 4.2 and 4.3

% Loday and anti-Loday accents (?)
\makeatletter
\newcommand{\oset}[3][0ex]{%
  \mathrel{\mathop{#3}\limits^{
    \vbox to#1{\kern-2\ex@
    \hbox{$\scriptstyle#2$}\vss}}}}
%
\newcommand{\uset}[3][0ex]{%
  \mathrel{\mathop{#3}\limits_{
    \vbox to#1{\kern-7\ex@
    \hbox{$\scriptstyle#2$}\vss}}}}
\makeatother
%
\newcommand{\loday}[1]{\smash{\overset{\frown}{#1}}}
\newcommand{\antiloday}[1]{\smash{\overset{\smile}{#1}}}
\newcommand{\upArc}[1]{\smash{\raisebox{.05cm}{$\uset[0ex]{#1}{\frown}$}}}
\newcommand{\downArc}[1]{\smash{\raisebox{-.05cm}{$\oset[.2ex]{#1}{\smile}$}}}
% yin and yang accents (?)
\newcommand{\yang}[1]{\smash{\overset{\backsim}{#1}}}
\newcommand{\yin}[1]{\smash{\overset{\sim}{#1}}}
% weak and strong equivalence on S_n
\newcommand{
\weakeq}{\asymp}
%\makeatletter
%\DeclareRobustCommand{\strongeq}{%
%  \mathbin{\mathpalette\scaletoplus@{\mkern1mu \smash{\begin{smallmatrix} \backsim \\[-.3cm] \sim \end{smallmatrix}} \mkern1mu}}%
%}
%\newcommand{\scaletoplus@}[2]{%
%  \vcenter{%
%    \hbox{\sbox\z@{$\m@th#1+$}\resizebox{\wd\z@}{!}{$\m@th#1#2$}}%
%  }%
%}
%\makeatother
\newcommand{\strongeq}{\mathbin{\smash{\begin{smallmatrix} \backsim \\[-.3cm] \sim \end{smallmatrix}}}}% % J: I would like to stack \backsim and \sim instead

@@

@@

Therefore, the quotientope is realized by a Minkowski sum of the shard polytopes defined by two families of arcs: the \defn{up arcs}~$\upArc{I}$ and the \defn{down arcs}~$\downArc{I}$, that go respectively above and below all elements of an interval $I$ of $[n]$.

@@

We call the congruence $\strongeq$ the \defn{strong rectangulation congruence}.
The \defn{strong rectangulotope} $\SRP(n)$ is the quotientope of the strong rectangulation congruence $\strongeq$.

@@

We classify the arcs into two families, depending on whether the arcs cross from above or below.
A \defn{yin arc} is defined by two consecutive nonempty intervals $I$ and $J$ of $[n]$ such that the arc goes above all elements of $I$ and below all elements of $J$.
A \defn{yang arc} is defined similarly, except that the arc goes below the elements of $I$ and above the elements of $J$.
Note that all up arcs and down arcs are both yin arcs and yang arcs.

@@

\begin{lemma}
  \label{lem:yinsp}
  Given a yin arc $A=(I,J)$, the yin shard polytope $\SP(A)$ is defined as the convex hull of the points
  $\{\b{0}\} \cup \set{\b{e}_i - \b{e}_j}{(i,j)\in I\times J}$.
\end{lemma}

\begin{lemma}
  \label{lem:yangsp}
  Given a yang arc $A=(I,J)$, the yang shard polytope $\SP(A)$ is a translate of the convex hull of the points
  $\{\b{0}\} \cup \set{\b{e}_j - \b{e}_i}{(i,j)\in I\times J}$.
\end{lemma}

Note that yin shard polytopes are translates of the Loday shard polytopes when $|J|=1$, and of the anti-Loday shard polytopes when $|I|=1$.
Similarly, Loday and anti-Loday shard polytopes are translates of some yang shard polytopes.

@@

\begin{lemma}
  \label{lem:strongMinkowski}
  The strong rectangulotope $\SRP(n)$ is realized by the Minkowski sum of all shard polytopes of yin and yang arcs.
\end{lemma}

Since the combinatorial type of a Minkowski sum is invariant to translation of the summands, we can use the polytopes defined in \cref{lem:yinsp,lem:yangsp} as a definition of the yin and yang shard polytopes.
Also note that since Loday and anti-Loday shard polytopes are, up to a translation, both yin and yang, each Loday and anti-Loday shard polytope is summed twice in the proposed realization.
This, again, does not change the combinatorial type of the polytope.

\cref{lem:strongsum} gives us an alternative description of $\SRP(n)$ as the sum of two opposite quotientopes defined by the \defn{yin} and \defn{yang congruences} respectively, where the yin congruence is defined by the arc ideal consisting of all yin arcs, and the yang congruence is defined by the yang arcs.

@@

\jean{That seems to work.}

>>>>

% Loday and anti-Loday accents (?)
\makeatletter
\newcommand{\oset}[3][0ex]{%
  \mathrel{\mathop{#3}\limits^{
    \vbox to#1{\kern-2\ex@
    \hbox{$\scriptstyle#2$}\vss}}}}
%
\newcommand{\uset}[3][0ex]{%
  \mathrel{\mathop{#3}\limits_{
    \vbox to#1{\kern-7\ex@
    \hbox{$\scriptstyle#2$}\vss}}}}
\makeatother
%
\newcommand{\loday}[1]{\smash{\overset{\frown}{#1}}}
\newcommand{\antiloday}[1]{\smash{\overset{\smile}{#1}}}
\newcommand{\upArc}[1]{\smash{\raisebox{.05cm}{$\uset[0ex]{#1}{\frown}$}}}
\newcommand{\downArc}[1]{\smash{\raisebox{-.05cm}{$\oset[.2ex]{#1}{\smile}$}}}
% yin and yang accents (?)
\newcommand{\yin}[1]{\smash{\overset{\sim}{#1}}}
\newcommand{\yang}[1]{\smash{\overset{\backsim}{#1}}}
\newcommand{\yinArc}[2]{\smash{\raisebox{.05cm}{$\uset[0ex]{#1}{\frown}$}\hspace{-.3ex}\raisebox{-.03cm}{$\oset[.2ex]{#2}{\smile}$}}}
\newcommand{\yangArc}[2]{\smash{\raisebox{-.03cm}{$\oset[.2ex]{#1}{\smile}$}\hspace{-.3ex}\raisebox{.05cm}{$\uset[0ex]{#2}{\frown}$}}}
% weak and strong equivalence on S_n
\newcommand{\weakeq}{\asymp}
\newcommand{\strongeq}{\mathbin{\smash{\begin{smallmatrix} \backsim \\[-.3cm] \sim \end{smallmatrix}}}}% % J: I would like to stack \backsim and \sim instead

@@

These arcs can be classified into two families of arcs: the up arcs and the down arcs.
For an interval $I$ of $[n]$ of size at least~$2$, the \defn{up arc}~$\upArc{I}$ (resp.~\defn{down arc}~$\downArc{I}$) is the arc that starts at~$\min(I)$ and ends at~$\max(I)$, and passes above (resp.~below) all remaining elements of~$I$, that is
\begin{align*}
\upArc{I} & \eqdef \big( \min(I), \max(I), I \ssm \{\min(I), \max(I)\}, \varnothing \big) \\
\qquad\text{and}\qquad
\downArc{I} & \eqdef \big( \min(I), \max(I), \varnothing, I \ssm \{\min(I), \max(I)\} \big).
\end{align*}
Note that the basic arc~$(i, i+1, \varnothing, \varnothing)$ is both an up and down arc.

@@

Therefore, the quotientope is realized by a Minkowski sum of the shard polytopes of the up arcs and the down arcs.
These quotientopes are very basic ones.

@@

We call the congruence $\strongeq$ the \defn{strong rectangulation congruence}.
The \defn{strong rectangulotope} $\SRP(n)$ is the quotientope of the strong rectangulation congruence~${\strongeq}$.

@@

We classify the arcs into two families, depending on whether the arcs cross from above or below.
For two consecutive intervals~$I$ and~$J$ of~$[n]$, the \defn{yin arc}~$\yinArc{I}{J}$ (resp.~\defn{yang arc}~$\yangArc{I}{J}$) is the arc starts at~$\min(I)$ and ends at~$\max(J)$, and goes above (resp.~below) the remaining elements of $I$ and below (resp.~above) the remaining elements of $J$, that is,
\begin{align*}
\yinArc{I}{J} & \eqdef \big( \min(I), \max(J), I \ssm \{\min(I)\}, J \ssm \{\max(J)\} \big) \\
\qquad\text{and}\qquad
\yangArc{I}{J} & \eqdef \big( \min(I), \max(J), J \ssm \{\max(J)\}, I \ssm \{\min(I)\} \big).
\end{align*}
Note that all up arcs and down arcs are both yin arcs and yang arcs.

@@

\begin{lemma}
  \label{lem:yinsp}
  The yin shard polytope $\SP(\yinArc{I}{J})$ is $\conv \{\b{0}\} \cup \set{\b{e}_i - \b{e}_j}{i \in I, \; j \in J}$.
\end{lemma}

\begin{lemma}
  \label{lem:yangsp}
  The yang shard polytope $\SP(\yangArc{I}{J})$ is a translate of~$\conv \{\b{0}\} \cup \set{\b{e}_j - \b{e}_i}{i \in I,\; j \in J}$.
\end{lemma}

%Note that yin shard polytopes are translates of the up shard polytopes when $|J|=1$, and of the down shard polytopes when $|I|=1$.
%Similarly, up and down shard polytopes are translates of some yang shard polytopes.

@@

\begin{lemma}
  \label{lem:strongMinkowski}
  The strong rectangulotope $\SRP(n)$ is realized by the Minkowski sum of all yin and yang shard polytopes.
\end{lemma}

Since the combinatorial type of a Minkowski sum is invariant to translation of the summands, we can use the polytopes defined in \cref{lem:yinsp,lem:yangsp} as a definition of the yin and yang shard polytopes.
Also note that since the up and down shard polytopes are, up to a translation, both yin and yang, each up and down shard polytope is summed twice in the proposed realization.
This, again, does not change the combinatorial type of the polytope.

\cref{lem:strongMinkowski} gives us an alternative description of $\SRP(n)$ as the sum of two opposite quotientopes defined by the \defn{yin} and \defn{yang congruences} respectively, where the yin congruence is defined by the arc ideal consisting of all yin arcs, and the yang congruence is defined by the yang arcs.

@@
